---
title: "GO specific CC"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: false
    code_folding: hide
---

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
.libPaths(c("/data/rajewsky/home/skim/R/usr_lib_Seurat/", "/data/rajewsky/home/skim/R/usr_lib_v4/"))
#.libPaths()
unloadNamespace("mgcv")
unloadNamespace("Matrix")


library(ggplot2)
library(gridExtra)
library(ggrepel)
library(dplyr)
library(cowplot)
library(paletteer)
library(stringr)
library(scales)
```

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
cor_mat <- readRDS("/data/rajewsky/projects/cdr1as_ko_snRNA/codes_github/cdr1as/DGEm/WTN_batch_corr_nested_DGEm.rda")




res3_bp_up <- readRDS(file = paste0("/data/rajewsky/projects/cdr1as_ko_snRNA/codes_github/cdr1as/Results/", "GO_Analysis_","KOm7oe_KO_SPECIFIC_UP", "_CC.rda"))
res3_bp_up <- res3_bp_up[which(res3_bp_up$p.value_elim < 0.05 & res3_bp_up$Annotated < 1000 & res3_bp_up$Annotated > 30), ] 
res3_bp_down <- readRDS(file = paste0("/data/rajewsky/projects/cdr1as_ko_snRNA/codes_github/cdr1as/Results/", "GO_Analysis_","KOm7oe_KO_SPECIFIC_DOWN", "_CC.rda"))
res3_bp_down <- res3_bp_down[which(res3_bp_down$p.value_elim < 0.05 & res3_bp_down$Annotated < 1000 & res3_bp_down$Annotated > 30), ] 


res4_bp_up <- readRDS(file = paste0("/data/rajewsky/projects/cdr1as_ko_snRNA/codes_github/cdr1as/Results/", "GO_Analysis_","KOm7oe_WT_SPECIFIC_UP", "_CC.rda"))
res4_bp_up <- res4_bp_up[which(res4_bp_up$p.value_elim < 0.05 & res4_bp_up$Annotated < 1000 & res4_bp_up$Annotated > 30), ] 
res4_bp_down <- readRDS(file = paste0("/data/rajewsky/projects/cdr1as_ko_snRNA/codes_github/cdr1as/Results/", "GO_Analysis_","KOm7oe_WT_SPECIFIC_DOWN", "_CC.rda"))
res4_bp_down <- res4_bp_down[which(res4_bp_down$p.value_elim < 0.05 & res4_bp_down$Annotated < 1000 & res4_bp_down$Annotated > 30), ] 


res1.df <- readRDS(file = "/data/rajewsky/projects/cdr1as_ko_snRNA/codes_github/cdr1as/Results/DESeq2_WTN_batch_corrected_res1.rda")
res2.df <- readRDS(file = "/data/rajewsky/projects/cdr1as_ko_snRNA/codes_github/cdr1as/Results/DESeq2_WTN_nested_corrected_res2.rda")
res3.df <- readRDS(file = "/data/rajewsky/projects/cdr1as_ko_snRNA/codes_github/cdr1as/Results/DESeq2_WTN_nested_corrected_res3.rda")
res4.df <- readRDS(file = "/data/rajewsky/projects/cdr1as_ko_snRNA/codes_github/cdr1as/Results/DESeq2_WTN_batch_corrected_res4.rda")

# up-genes
set1 <- res1.df[which(res1.df$log2FoldChange > 0.5 &  res1.df$padj < 0.05), ]
set2 <- res2.df[which(res2.df$log2FoldChange > 0.5 &  res2.df$padj < 0.05), ]
set3 <- res3.df[which(res3.df$log2FoldChange > 0.5 &  res3.df$padj < 0.05), ]
set4 <- res4.df[which(res4.df$log2FoldChange > 0.5 &  res4.df$padj < 0.05), ] 

genes1 <- c(set1, set2, set3, set4) %>% unique()

# down-genes
set5 <- res1.df[which(res1.df$log2FoldChange < -0.5 &  res1.df$padj < 0.05), ]
set6 <- res2.df[which(res2.df$log2FoldChange < -0.5 &  res2.df$padj < 0.05), ]
set7 <- res3.df[which(res3.df$log2FoldChange < -0.5 &  res3.df$padj < 0.05), ]
set8 <- res4.df[which(res4.df$log2FoldChange < -0.5 &  res4.df$padj < 0.05), ]


# res3 specific
specific_genes_up3 <- Reduce(setdiff, list(set3$gene,set1$gene,set2$gene,set4$gene)) 
specific_genes_down3 <- Reduce(setdiff, list(set7$gene,set5$gene,set6$gene,set8$gene)) 

# res4 specific
specific_genes_up4 <- Reduce(setdiff, list(set4$gene,set1$gene,set2$gene,set3$gene)) 
specific_genes_down4 <- Reduce(setdiff, list(set8$gene,set5$gene,set6$gene,set7$gene)) 

```

## 1. Results

### res3 specific-up

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
DT::datatable(res3_bp_up)
```

### res3 specific-down

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
DT::datatable(res3_bp_down)
```

### res4 specific-up

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
DT::datatable(res4_bp_up)
```

### res4 specific-down

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
DT::datatable(res4_bp_down)
```

## 2. Term check 

### 2-1. extracellular  res3 specific-up

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res3_bp_up[grepl("extracellular ", res3_bp_up$Term), ]
```

### 2-2. extracellular res4 specific-up

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res4_bp_up[grepl("extracellular", res4_bp_up$Term), ]
```

### 2-3. extracellular res3 specific-down

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res3_bp_down[grepl("G protein", res3_bp_down$Term), ]
```

### 2-4. extracellular res4 specific-down

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res4_bp_down[grepl("extracellular", res4_bp_down$Term), ]
```

### 3-1. synap  res3 specific-up

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res3_bp_up[grepl("synap", res3_bp_up$Term), ]
```

### 3-2. synap res4 specific-up

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res4_bp_up[grepl("synap", res4_bp_up$Term), ]
```

### 3-3. synap res3 specific-down

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res3_bp_down[grepl("synap", res3_bp_down$Term), ]
```

### 3-4. synap res4 specific-down

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res4_bp_down[grepl("synap", res4_bp_down$Term), ]
```

### 4-1. glutamat res3 specific-up

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res3_bp_up[grepl("glutamat", res3_bp_up$Term), ]
```

### 4-2. glutamat res4 specific-up

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res4_bp_up[grepl("glutamat", res4_bp_up$Term), ]
```

### 4-3. glutamat res3 specific-down

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res3_bp_down[grepl("glutamat", res3_bp_down$Term), ]
```

### 4-4. glutamat res4 specific-down

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res4_bp_down[grepl("glutamat", res4_bp_down$Term), ]
```

### 5-1. vesicle res3 specific-up

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res3_bp_up[grepl("vesicle", res3_bp_up$Term), ]
```

### 5-2. vesicle res4 specific-up

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res4_bp_up[grepl("vesicle", res4_bp_up$Term), ]
```

### 5-3. vesicle res3 specific-down

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res3_bp_down[grepl("vesicle", res3_bp_down$Term), ]
```

### 5-4. vesicle res4 specific-down

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res4_bp_down[grepl("vesicle", res4_bp_down$Term), ]
```

### 6-1. mitoch res3 specific-up

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res3_bp_up[grepl("mitoch", res3_bp_up$Term), ]
```

### 6-2. mitoch res4 specific-up

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res4_bp_up[grepl("mitoch", res4_bp_up$Term), ]
```

### 6-3. mitoch res3 specific-down

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res3_bp_down[grepl("mitoch", res3_bp_down$Term), ]
```

### 6-4. mitoch res4 specific-down

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res4_bp_down[grepl("mitoch", res4_bp_down$Term), ]
```

### 7-1. glutamate  res3 specific-up

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res3_bp_up[grepl("glutamate", res3_bp_up$Term), ]
```

### 7-2. glutamate res4 specific-up

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res4_bp_up[grepl("glutamate", res4_bp_up$Term), ]
```

### 7-3. glutamate res3 specific-down

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res3_bp_down[grepl("glutamate", res3_bp_down$Term), ]
```

### 7-4. glutamate res4 specific-down

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res4_bp_down[grepl("glutamate", res4_bp_down$Term), ]
```

### 8-1. endopla  res3 specific-up

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res3_bp_up[grepl("endopla", res3_bp_up$Term), ]
```

### 8-2. endopla res4 specific-up

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res4_bp_up[grepl("endopla", res4_bp_up$Term), ]
```

### 8-3. endopla res3 specific-down

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res3_bp_down[grepl("endopla", res3_bp_down$Term), ]
```

### 8-4. endopla res4 specific-down

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
res4_bp_down[grepl("endopla", res4_bp_down$Term), ]
```

## 3. Figures

### 3-1. Dot plot

## 3. Figures

### 3-1. Dot plot

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
Terms <- c("cortical cytoskeleton",
           "extracellular vesicle",
           "glutamatergic synapse",
           "excitatory synapse", 
           "GABA-ergic synapse",
           "synaptic vesicle", 
           "transport vesicle",
           "exocytic vesicle",
           "mitochondrial matrix"
           )

res3_bp_up_sub <- res3_bp_up[res3_bp_up$Term %in% Terms, ]
res3_bp_up_sub$exp <- "KOm7oe_vs_KO_Up"

res4_bp_up_sub <- res4_bp_up[res4_bp_up$Term %in% Terms, ]
res4_bp_up_sub$exp <- "KOm7oe_vs_WT_Up"

res3_bp_down_sub <- res3_bp_down[res3_bp_down$Term %in% Terms, ]
res3_bp_down_sub$exp <- "KOm7oe_vs_KO_Down"

res4_bp_down_sub <- res4_bp_down[res4_bp_down$Term %in% Terms, ]
res4_bp_down_sub$exp <- "KOm7oe_vs_WT_Down"

GO.df <- rbind(
           res3_bp_down_sub,
           res4_bp_down_sub,
          
           res3_bp_up_sub,
           res4_bp_up_sub)


GO.df <- GO.df %>% dplyr::select(Term, Annotated, Significant, p.value_elim, exp)

GO.df$ratio <- GO.df$Significant / GO.df$Annotated

GO.df$exp <- factor(GO.df$exp, levels = c(
                                          "KOm7oe_vs_KO_Up", "KOm7oe_vs_KO_Down",
                                          "KOm7oe_vs_WT_Up", "KOm7oe_vs_WT_Down"))

GO.df$p.value_elim <- as.numeric(GO.df$p.value_elim)
GO.df$p.value_elim <- -log10(GO.df$p.value_elim)
GO.df$Term <- factor(GO.df$Term, levels = rev(unique(GO.df$Term)))

GO.df$Term <- factor(GO.df$Term,
                     levels = rev(c(Terms[1],Terms[4],Terms[7],Terms[2],Terms[3],
                                    Terms[6],Terms[8], Terms[9],Terms[5])))



```

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=8}
g1 <- ggplot(GO.df, aes(x=exp,
               y=Term,
               colour=p.value_elim,
               size=ratio)) +
        geom_point() +
        expand_limits(x=0) +
        labs(x="Expected", y="GO term", colour="-log10(p-value)", size="gene ratio") +
  theme_minimal() + theme(axis.title = element_blank(),
                          axis.text.y =  element_text(size = 10, color = "black"),
                          axis.text.x = element_text(angle = 90, size = 10, vjust = 0.5, color = "black")) +
  scale_color_gradientn("-log10(p-value)", colours = c(paletteer_d("ggsci::amber_material"))) +
  guides(color = guide_colourbar(order = 1), size = guide_legend(order = 2))


g1

# ggsave(filename = "/data/rajewsky/home/skim/Neuronal_Activity_Cledi/201911_m7_overexpression/publication_codes_figures/GO_CC_specific_figures_dot_plot.pdf",
#        plot = g1,
#        scale = 1, width = 7, height = 7, units = "in", device = cairo_pdf,
#        dpi = 300)

```