---
title: "venn diagram genes 5SSKO"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: false
    code_folding: hide
---

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
.libPaths(c("/data/rajewsky/home/skim/R/usr_lib_Seurat/", "/data/rajewsky/home/skim/R/usr_lib_v4/"))
#.libPaths()
unloadNamespace("mgcv")
unloadNamespace("Matrix")

library(WGCNA)
library(flashClust)
library(DESeq2)
library(genefilter)
library(ggplot2)
library(plotly)
library(gridExtra)
library(ggrepel)
library(dplyr)
library(cowplot)
library(DT)
library(bnlearn)
library(Rgraphviz)
library(gplots)
library(DT)
library(GGally)
library(penalized)
library(visNetwork)
#BiocManager::install("clusterProfiler", lib = "/data/rajewsky/home/skim/R/usr_lib_v4/")
#library(clusterProfiler)
library(pcaExplorer)
#BiocManager::install("pcaExplorer", lib = "/data/rajewsky/home/skim/R/usr_lib_v4/")
library(topGO)
library(igraph)
library(paletteer)
library(stringr)
library(devtools)
#library(limmaDE2)
library(scales)
library(RColorBrewer)
library(VennDiagram)
library(ggvenn)
library(tidyr)
library(readxl)
```

```{r}
library(clusterProfiler)

#########################
# informative data load
#########################

DGEm_normalized <- readRDS(file = "/data/rajewsky/home/skim/Neuronal_Activity_Cledi/201911_m7_overexpression/publication_codes_figures/DGE_matrix_normalized_nested.rda")

# miR-7 target genes 
mir7target.df <- read.table("/data/rajewsky/home/skim/Neuronal_Activity_Cledi/201911_m7_overexpression/R_Scripts/external_files/TargetScan7.2__miR-7-5p.predicted_targets_lists_without_header.txt")

# load pheno genes
cledi_pheno_genes <- read_excel("/data/rajewsky/home/skim/Neuronal_Activity_Cledi/201911_m7_overexpression/R_Scripts/external_files_2nd/biased_list_of_phenotype_related_genes.xlsx", col_names = FALSE)

# TF genes
TF_list <- read.csv("/data/rajewsky/projects/cdr1as_ko_snRNA/3rd_sequencing_run/pyscenic/Mouse_TFs_list.txt", stringsAsFactors = FALSE)
TF_list[725,1] <- "Auts2"
colnames(TF_list) <- "TF"

#########################
# DE data load
#########################

res1.df <- readRDS(file = "/data/rajewsky/projects/cdr1as_ko_snRNA/codes_github/cdr1as/Results/DESeq2_WTJ_5SS_res1.rda")
res2.df <- readRDS(file = "/data/rajewsky/projects/cdr1as_ko_snRNA/codes_github/cdr1as/Results/DESeq2_WTJ_5SS_nested_corrected_res2.rda")
res3.df <- readRDS(file = "/data/rajewsky/projects/cdr1as_ko_snRNA/codes_github/cdr1as/Results/DESeq2_WTJ_5SS_nested_corrected_res3.rda")
res4.df <- readRDS(file = "/data/rajewsky/projects/cdr1as_ko_snRNA/codes_github/cdr1as/Results/DESeq2_WTJ_5SS_res4.rda")


# up-genes
set1 <- res1.df[which(res1.df$log2FoldChange > 0.5 &  res1.df$padj < 0.05), ]$gene
set2 <- res2.df[which(res2.df$log2FoldChange > 0.5 &  res2.df$padj < 0.05), ]$gene
set3 <- res3.df[which(res3.df$log2FoldChange > 0.5 &  res3.df$padj < 0.05), ]$gene
set4 <- res4.df[which(res4.df$log2FoldChange > 0.5 &  res4.df$padj < 0.05), ]$gene 

genes1 <- c(set1, set2, set3, set4) %>% unique()

# down-genes
set5 <- res1.df[which(res1.df$log2FoldChange < -0.5 &  res1.df$padj < 0.05), ]$gene
set6 <- res2.df[which(res2.df$log2FoldChange < -0.5 &  res2.df$padj < 0.05), ]$gene
set7 <- res3.df[which(res3.df$log2FoldChange < -0.5 &  res3.df$padj < 0.05), ]$gene
set8 <- res4.df[which(res4.df$log2FoldChange < -0.5 &  res4.df$padj < 0.05), ]$gene 

genes2 <- c(set5, set6, set7, set8) %>% unique()

# list of DE genes by experiment
Cdr1as_DE_genes <- c(set1, set5)
WT_miR7_DE_genes <- c(set2, set6)
KO_miR7_DE_genes <- c(set3, set7)
KO_WT_miR7_DE_genes <- c(set4, set8)


all_genes <- c(genes1, genes2) %>% unique()
# genes exist in both up- and down-
intersect(genes1, genes2)

genes.df <- data.frame(gene = all_genes)
         
```

## 1. up

```{r warning=FALSE, message=FALSE, fig.height=15, fig.width=20}
x <- list(
  KO_vs_WT = res1.df[which(res1.df$log2FoldChange > 0.5 &  res1.df$padj < 0.05), ]$gene,
  WTm7oe_vs_WT = res2.df[which(res2.df$log2FoldChange > 0.5 &  res2.df$padj < 0.05), ]$gene,
  KOm7oe_vs_KO = res3.df[which(res3.df$log2FoldChange > 0.5 &  res3.df$padj < 0.05), ]$gene,
  KOm7oe_vs_WT = res4.df[which(res4.df$log2FoldChange > 0.5 &  res4.df$padj < 0.05), ]$gene
  )


#x <- list(set1, set2, set3, set4)

g1 <- ggvenn(
  x, 
  stroke_linetype = 1, 
  fill_color = c("#0073C2FF", "#EFC000FF", "#868686FF", "#CD534CFF"),
  stroke_size = 1, set_name_size = 9, text_size = 12
  ) + 
  theme(plot.margin=unit(c(0,0,0,0),"cm"))

# ggsave(filename = "/data/rajewsky/home/skim/Neuronal_Activity_Cledi/201911_m7_overexpression/publication_codes_figures/venn_diagram_up_all.pdf",
#        plot = g1,
#        scale = 1, width = 20, height = 15, units = "in", device = cairo_pdf,
#        dpi = 300)

g1
```

### all expressed 

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
Reduce(intersect, list(set1,set2,set3, set4)) 
```

### specific KO vs WT

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
Reduce(setdiff, list(set1,set2,set3, set4)) 
```

### specific WTm7oe vs WT

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
Reduce(setdiff, list(set2,set1,set3, set4)) 
```

### specific KOm7oe vs KO

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
Reduce(setdiff, list(set3,set1,set2, set4)) 
```

### specific KOm7oe vs WT

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
Reduce(setdiff, list(set4,set1,set2, set3)) 
```

### co-exp KOm7oe vs KO & KOm7oe vs WT

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
list1 <- Reduce(intersect, list(set4, set3))
list2 <- Reduce(intersect, list(set4, set2))
list3 <- Reduce(intersect, list(set4, set1))

list <- list1[!list1 %in% c(list2, list3)]
list
```

### co-exp WTm7oe vs WT & KOm7oe vs WT

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
list1 <- Reduce(intersect, list(set4, set3))
list2 <- Reduce(intersect, list(set4, set2))
list3 <- Reduce(intersect, list(set4, set1))

list <- list2[!list2 %in% c(list1, list3)]
list
```

### co-exp WTm7oe vs WT & KOm7oe vs WT & KOm7oe vs KO

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
list1 <- Reduce(intersect, list(set4, set3, set2))
list2 <- Reduce(intersect, list(set4, set1))

list <- list1[!list1 %in% c(list2)]
list
```

### co-exp specific KO vs WT & KOm7oe vs WT

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
list1 <- Reduce(intersect, list(set4, set1))
list2 <- Reduce(intersect, list(set4, set2, set1))
list <- list1[!list1 %in% c(list2)]
list
```

### co-exp specific KO vs WT & KOm7oe vs WT & KO vs WT

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
list1 <- Reduce(intersect, list(set4, set2, set1))
list2 <- Reduce(intersect, list(set3, set2, set1))
list <- list1[!list1 %in% c(list2)]
list
```

### co-exp specific KOm7oe vs KO & WTm7oe vs WT 

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
list1 <- Reduce(intersect, list(set3, set2))
#list2 <- Reduce(intersect, list(set4, set1))
list <- list1[!list1 %in% c(set4, set1)]
list
```


## 2. down

```{r warning=FALSE, message=FALSE, fig.height=15, fig.width=20}
x <- list(
  KO_vs_WT = res1.df[which(res1.df$log2FoldChange < -0.5 &  res1.df$padj < 0.05), ]$gene,
  WTm7oe_vs_WT = res2.df[which(res2.df$log2FoldChange < -0.5 &  res2.df$padj < 0.05), ]$gene,
  KOm7oe_vs_KO = res3.df[which(res3.df$log2FoldChange < -0.5 &  res3.df$padj < 0.05), ]$gene,
  KOm7oe_vs_WT = res4.df[which(res4.df$log2FoldChange < -0.5 &  res4.df$padj < 0.05), ]$gene
  )


#x <- list(set1, set2, set3, set4)

g1 <- ggvenn(
  x, 
  stroke_linetype = 1, 
  fill_color = c("#0073C2FF", "#EFC000FF", "#868686FF", "#CD534CFF"),
  stroke_size = 1, set_name_size = 9, text_size = 12
  ) + 
  theme(plot.margin=unit(c(0,0,0,0),"cm"))
g1
# ggsave(filename = "/data/rajewsky/home/skim/Neuronal_Activity_Cledi/201911_m7_overexpression/publication_codes_figures/venn_diagram_up_all.pdf",
#        plot = g1,
#        scale = 1, width = 20, height = 15, units = "in", device = cairo_pdf,
#        dpi = 300)

```

### specific KO vs WT

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
Reduce(setdiff, list(set5,set6,set7, set8)) 
```

### specific WTm7oe vs WT

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
Reduce(setdiff, list(set6,set5,set7, set8)) 
```

### specific KOm7oe vs KO

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
Reduce(setdiff, list(set7,set5,set6, set8)) 
```

### specific KOm7oe vs WT

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
Reduce(setdiff, list(set8,set5,set6, set7)) 
```

### co-exp KOm7oe vs KO & KOm7oe vs WT

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
list1 <- Reduce(intersect, list(set8, set7))
list2 <- Reduce(intersect, list(set8, set6))
list3 <- Reduce(intersect, list(set8, set5))

list <- list1[!list1 %in% c(list2, list3)]
list
```

### co-exp WTm7oe vs WT & KOm7oe vs WT & KOm7oe vs KO

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
list1 <- Reduce(intersect, list(set8, set7, set6))
list2 <- Reduce(intersect, list(set8, set5))

list <- list1[!list1 %in% c(list2)]
list
```

### co-exp WTm7oe vs WT & KOm7oe vs WT

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
list1 <- Reduce(intersect, list(set8, set6, set5))
list <- list1[!list1 %in% c(set7)]
list
```

### co-exp WTm7oe vs WT & KOm7oe vs WT

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
list1 <- Reduce(intersect, list(set8, set6))
list <- list1[!list1 %in% c(set5, set7)]
list
```

### co-exp specific KO vs WT & KOm7oe vs WT

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
list1 <- Reduce(intersect, list(set8, set5))
list2 <- Reduce(intersect, list(set8, set6, set5))
list <- list1[!list1 %in% c(list2)]
list
```

### co-exp specific KOm7oe vs KO & WTm7oe vs WT 

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
list1 <- Reduce(intersect, list(set7, set6))
#list2 <- Reduce(intersect, list(set4, set1))
list <- list1[!list1 %in% c(set8, set5)]
list
```

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}

```

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
sessionInfo()
```

